---
title: "Regression Models Course Project"
author: "Youness BOUTYOUR"
date: "October 07, 2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
[Link to project on GitHUB](https://github.com/nicedev2020/Regression-Models-Course-Project)

[Link to project on RPub](https://rpubs.com/nicedev/RegressionModelsCourseProject)

# Executive summary

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:      
* "Is an automatic or manual transmission better for MPG"     
* "Quantify the MPG difference between automatic and manual transmissions"     

# Analysis

## Exploratory analysis

```{r , warning=FALSE, message = FALSE, echo=TRUE}
library (datasets)
data(mtcars) # Loading data
head(mtcars) # Dataset's head
dim(mtcars) # Row's numbers and variable's quantity 
```

Let's test hypotesys what automatic and manual transmission are the same on average for MPG?
```{r}
result <- t.test(mtcars$mpg ~ mtcars$am) 
result$p.value
``` 

Since the p-value is 0.00137, we reject our null hypothesis. So, the automatic and manual transmissions are from different populations. Let's show difference:

```{r}
result$estimate
``` 

```{r , warning=FALSE, message = FALSE, echo=TRUE}
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- as.factor(mtcars$am)
boxplot(mpg ~ am, 
        data = mtcars, 
        ylab = "Miles Per Gallon (MPG)", 
        xlab = "Transmission Type",
        col = (c("pink","blue")))
```

## Regression analysis

```{r, warning=FALSE, message = FALSE, echo=TRUE}
fit_SLR <- lm(mpg ~ factor(am), data=mtcars)
summary(fit_SLR)
```  

The adjusted R squared value is only 33.8% of the regression variance can be explained by our model. 
Let's see how will other predictor variables impact.

```{r, warning=FALSE, message = FALSE, echo=TRUE}
data(mtcars)
fit_MLR <- lm(mpg ~ . ,data=mtcars)
summary(fit_MLR)
```

```{r, warning=FALSE, message = FALSE, echo=TRUE}
cor(mtcars)[1,]
```
From the  output we can see cyl,wt,hp,disp show strong correlations and significance for the model. Hence we choose those variables plus am for a linear model. This gives us the following model below:

```{r, warning=FALSE, message = FALSE, echo=TRUE}
fit_MLR_adjusted <- lm(mpg ~ wt+hp+disp+cyl+am, data = mtcars)
summary(fit_MLR_adjusted)
```

## Residual Analysis and Diagnostics

```{r, warning=FALSE, message = FALSE, echo=TRUE}
par(mfrow = c(2, 2))
plot(fit_MLR_adjusted)
```

According to the residual plots:  
1. The Residuals vs. Fitted plot shows no consistent pattern, supporting the accuracy of the independence assumption.  
2. The Normal Q-Q plot indicates that the residuals are normally distributed because the points lie closely to the line.  
3. The Scale-Location plot confirms the constant variance assumption, as the points are randomly distributed.  
4. The Residuals vs. Leverage argues that no outliers are present, as all values fall well within the 0.5 bands.  

# Conclusions 

Using the final multivariable regression model put together we can see the multiple R squared value is much higher at 0.83, where 83% of the regression variance can be explained by the chosen variables. We can thus conclude that wt, hp, disp and cyl are confounding variables in the relationship between 'am and 'mpg' and that manual transmission cars on average have 1.55 miles per gallon more than automatic cars.
